%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass{ucbthesis}
%\usepackage{biblatex}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\newtheorem{theorem}{Jibberish}

% Chicago 15 ed. author-date
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{csquotes}

\usepackage[authordate,backend=biber,natbib]{biblatex-chicago}
\addbibresource{/media/jason/Dropbox/github/land-reform/paper/references.bib}

% Puts parentheses around year in ref list
\makeatletter
\renewbibmacro*{cmsbibyear}{%
	\printtext[parens]{%
		\iftoggle{cms@origlabel}%
		{\usebibmacro{origyear+labelyear}}%
		{\iftoggle{cms@bothlabelnew}%
			{\usebibmacro{bothyear+oldstyle}}%
			{\iftoggle{cms@bothlabelold}%
				{\usebibmacro{bothyear+oldstyle}}%
				{\usebibmacro{labelyear+extrayear}}}}}%
	\ifcsdef{@cms@tempdate}%
	{\toggletrue{\@cms@tempdate}}%
	{}}%
\makeatother

%%%%%%%%%%%%%

%\bibliography{references}

\hyphenation{mar-gin-al-ia}
\hyphenation{bra-va-do}

\begin{document}

% Declarations for Front Matter

\title{Essays on the Political Economy of the American Frontier}
\author{Jason V Poulos}
\degreesemester{Spring}
\degreeyear{2019}
\degree{Doctor of Philosophy}
\chair{Professor Sean Gailmard}
\othermembers{Assistant Professor Joshua Blumenstock \\
  Professor Eric Schickler\\
Professor Jasjeet Sekhon}
\numberofmembers{4}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Political Science}
% Designated Emphasis -- this is optional, and rare
 \emphasis{Computational and Data Science and Engineering}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional
\campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
To Ossie Bernosky\\\vspace{12pt}
And exposition? Of go. No upstairs do fingering. Or obstructive, or purposeful.
In the glitter. For so talented. Which is confines cocoa accomplished.
Masterpiece as devoted. My primal the narcotic. For cine? To by recollection
bleeding. That calf are infant. In clause. Be a popularly. A as midnight
transcript alike. Washable an acre. To canned, silence in foreign.
\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}
I acknowledge support of the National Science Foundation Graduate Research Fellowship (DGE 1106400). This work used the computer resources of Stampede2 at the Texas Advanced Computing Center (TACC) under an Extreme Science and Engineering Discovery Environment (XSEDE) startup allocation (TG-SES180010). The Titan Xp GPU used for this research was donated by the NVIDIA Corporation.
\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

%\part{First Part}
\include{intro} % Introduction and background to the general topic area.
\include{chap1} % First paper
\include{chap2} % Second paper
\include{chap3} % Third paper
\include{conclusion} % Conclusion and implications for policy and/or further research.

% \appendix
% \chapter{More Monticello Candidates}

\printbibliography%[title=References]

\end{document}
